Dark matter haloes: a multistream view

### 2018-03-07 Draft/Chapter5/Paper2017b_Nobmath_old.tex
initial commit

\hl{The web-like distribution of matter initially revealed by redshift surveys (with less than 300 galaxies by} \citealt{Gregory1978} \hl{and around 1000 galaxies by} \citealt{DeLapparent1986}) \hl{and numerical modeling (using N-body simulations of around 30000 particles by} \citealt{Shandarin1983} and \citealt{Klypin1983a}) \hl{pioneered morphological investigations of the cosmic web structures (see} \citealt{Bond1996}, also reviews by \citealt{Shandarin1989} and \citealt{Weygaert2008}). \hl{Detailed mapping of the Universe has crossed three million objects today, by catalogues such as the Sloan Digital Sky Survey (SDSS;} \citealt{Albareti2016}). The upcoming Large Synoptic Survey Telescope (LSST; \citealt{lsst2009}) \hl{is expected to probe the nature of dark matter using several billion galaxies. On the other hand, cosmological simulations have improved immensely in several aspects -- numerical techniques, parallelization schemes, inclusion of various physical processes, volume and resolution (some of these developments are summarized in }\citealt{Bertschinger1998} and \citealt{Bagla1997}). Modern state-of-the-art simulations like the Illustris Project \cite{Vogelsberger2014} and Q-Continuum \cite{Heitmann2015} \hl{use more than a billion dark matter particles. Finally, the ever improving data analysis techniques have resulted in new and sophisticated density estimators, geometrical and topological indicators. A plethora of algorithms for identifying and characterizing dark matter structures have emerged in last two decades (a summary on cosmological data analysis is highlighted in} \citealt{Weygaert2009a}). \hl{Considering all these improvements, it is worth noting that the proto-structures detected in the modern simulations are qualitatively similar to the quasi-linear description of clustering by Zeldovich Approximation (ZA;} \citealt{Zeldovich1970}). \hl{Location and properties of these structures, i.e., the voids, walls, filaments and haloes maybe inconsistent across different structure finding algorithms, but that is primarily due to varied definitions. }

@@

Most of structure finders are halo finders only and majority of them are stemmed from three underlying algorithms. One of them is the SO (Spherical Over-density) halo finder that defines halos  as spherical regions whose mass density exceeds the mean density by a specified factor \citep{Press1974}. 
Another is the FOF  (Friends-of-Friends) halo finder describing haloes  as the groups of particles separated less than a specified linking length often chosen as 0.2 times the mean particle separation \citep{Davis1985}. 
\hl{ The FOF can be also used for identifying filaments and walls/pancakes by increasing the linking length} (\citealt{Zeldovich1982}, \citealt{Shandarin1983a}, \citealt{Shandarin2010b} )  %(refs in tex file and in additional\_refs.bib)}. 
%\citep{Zeldovich_etal:82,Shandarin:83,Shandarin_etal:10 } 
Finally the DENMAX (DENsity MAXimum) halo finder assumes that the halos are the peaks of the density  fields and thus selects the particles concentrated in the vicinity of the density maxima \citep{Bertschinger1991}. One of the common features of these  techniques is that all three are based on density, in one form or another. And all of them depend on free parameters that are chosen chiefly on the `merits principle' \citep{Forero-Romero2009a} rather than on physics. Over the years all three kinds of the halo finders have been experiencing various modifications and improvements.  A few examples from a long list of these modifications may include:

@@

 \hl{ A detailed comparisons of several halo/sub-halo finders is provided in }
%done in projects like 
\cite{Knebe2011a}, \cite{Knebe2013}, \cite{Onions2012}.\hl{ In a nice summary discussing  these developments as well as describing a few new suggestions they  concluded that there was no general consensus for a precise definition of a halo or a sub-halo. Consequently, there were different estimates of number of haloes, halo mass functions, halo centre locations, boundaries and other parameters. }
%A  are given in comparison project papers quoted above.

@@

\hl{For FOF, the free parameter of linking length is generally taken as $b = 0.2$ times the mean separation of particles at $z=0$. This inter-particle separation corresponds to $\Delta_{vir} \approx 180$ if the halo has an isothermal density profile, $\rho \propto r^{-2}$. Using percolation theory,} \cite{More2011} \hl{argued that this corresponds to a rather wide range of over-densities depending on halo mass and density profiles. They found out that $b = 0.2$ corresponds to local over-density $\delta$ within the enclosed halo to be in the range of 250 to 600. Moreover, the resulting FOF-haloes need not have a compact geometry: often the haloes are irregularly shaped, which is unlikely if the halos are virialized.
Hence modern algorithms re-define the halo boundaries by excluding particles using post-processing techniques. In recent simulations with clear delineation of walls and filaments} \cite{Angulo2013a}, \hl{$b=0.05$ was used for finding FOF-haloes since the traditional value of $b = 0.2$ corresponded to structures that percolate into the web structure. }

\hl{Absence of dynamical traits in the FOF and SO algorithms are arguably more crucial. In phase-space, the halo collapse models show collisionless DM particles in oscillatory motions about a core, at successive foldings of the phase-space sheet. The velocity field within each oscillatory spiral is multi-valued in physical space. Incrementing {\it multistream} shells, separated by {\it caustic} surfaces sequentially trace the structures of the cosmic web - walls, filaments and the haloes. Majority of the mass accretion into the haloes along the filaments: from lower multistreams into higher. Thus the DM haloes are not independent of filaments around them, and the hierarchical layers of multistreams represents this portrait precisely. This picture of structure formation was initially theorized using ZA} \cite{Zeldovich1970} and in context of caustics \cite{Arnold1982} as well as in the Adhesion Approximation (\citealt{Gurbatov1989}, \citealt{Kofman1992}). 
\cite{Shandarin1989} reviewed gravitational evolution of density perturbations in the context.

\hl{It has been demonstrated that the multistream field in Eulerian space can be computed directly from the Lagrangian sub-manifold} (\citealt{Shandarin2012} and \citealt{Abel2012b}). About 90\% of the field is single-streaming voids, and the rest of the volume comprises of multistream walls, filaments and haloes. \cite{Ramachandra2015} found the multistream value of $n_{str} \approx 90$ corresponds to virial density $\Delta_{vir} = 200$. On the other hand, DM particles are identified by \citep{Falck2012} as belonging to haloes if they undergo flip-flop along 3 orthogonal axes. \hl{These analyses have opened up a new avenue in studies of halo formation, both qualitative and quantitative. Re-investigations of halo spins, physical radii of the halo, sub-structure in the light of streaming phenomena have shown that the halo structures and formations are more complicated than previously believed.} \cite{Vogelsberger2011} \hl{investigated the distribution of streams in small haloes at various redshifts. They concluded that tracking caustics and streams is better than density, since density fields are noisy in the dense inner regions of haloes. In another study,} \cite{More2015} \hl{argued that the `splashback radius' - distance from the halo core to the first caustic enumerated from outside - is a better physical indicator of DM halo boundary than the virial radius (also see the discussion on turn-around radius of bound objects by} \citealt{Lee2016a}). \cite{Angulo2013a} \hl{also agree with the view that the locally overdense regions correspond better with the volumes within the first caustic than the virialized DM clumps. Recent toy model of anisotropic halo collapse by} \cite{Neyrinck2016} considers intersecting multistream filaments forming spinning nodes. \cite{Ramachandra2017} showed that the virial surfaces of FOF haloes have varying number of streams, including single-streams. Study by \cite{Shandarin2016} delineated the rich sub-structure of haloes using another derivative parameter from the Lagrangian sub-manifold called the `flip-flop' defined on the Lagrangian space.

@@

\hl{New section}

\begin{figure*} 
\centering\includegraphics[height=7cm]{fig1.pdf} 
\caption{Dynamical collapse of dark matter in one-dimensional universe: Top panels show the $(\bmath{p}, \bmath{x})$ phase-space manifold of the dark matter sheet at redshifts $z_1$, $z_2$, $z_3$ and $z = 0$. Dots represent the dark matter particles. The momentum values are chosen at arbitrary scales. Bottom panels show the corresponding multistream field multistream field $n_{str}(\bmath{x}, z)$ (red) and density field $\rho(\bmath{x}, z)$ (gray). }
\label{fig:1d}
\end{figure*}

We begin with a simple illustration showing the formation of a few haloes in a one-dimensional simulation. Dark matter clustering in a (1+1)-dimensional phase-space $(\bmath{p}, \bmath{x})$ (where $p$ is the momentum and $x$ is the co-moving Eulerian coordinate) at four successive time steps is shown in the top panels of \autoref{fig:1d}. The lower panels show the corresponding multistream field (\citealt{Shandarin2012} and \citealt{Abel2012b}) $n_{str}(\bmath{x}, z)$ (red) and density field $\rho(\bmath{x}, z)$ (gray). At $z_1$ (left-most panel), velocity is single-valued in Eulerian co-ordinates shown, except at a small three-stream region near $\bmath{x} = 5\pi/4$. This is the first instance of multistreaming in the region, which was previously had $n_{str} = 1$ throughout. The interface of $n_{str} = 1$ and $n_{str} = 3$ regions is also the location of the first caustic. On the other hand, the density calculated at a high resolution shows variations, even in the mono-streaming regions. The variations are especially more pronounced around  the caustic (near $\bmath{x} = 5\pi/4$).

@@

\begin{figure} 
\centering\includegraphics[width=8cm]{fig2.pdf} 
\caption{Multistream field contours: The multistream field is calculated at 8 times the native resolution. Each inner convex blobs (red) surround local multistream maxima inside. Surrounding outer shell(blue) is not convex throughout the surface, and the outermost gray multistream surface displays a filamental geometry.}
\label{fig:full}
\end{figure}

@@

\hl{The emphasis of this paper is to demonstrate the use of multistream field in identifying potential dark matter haloes, and not a full statistical analysis of halo properties. For this purpose, we have run simulations at two different mass resolutions (number of particles $N_p = 128^3$ and $256^3$, and respective mass of particles, $m_p = 3.65 \times 10^{10} h^{-1} M_{\sun}$ and $ 4.57 \times 10^{9} h^{-1} M_{\sun}$), with the same periodic side length $L = 100 h^{-1} Mpc$. The gravitational softening length $\epsilon = 20 h^{-1} kpc$ and $10 h^{-1} kpc$ for low and high resolution simulation respectively. The initial conditions are generated by {MUSIC}} \citep{Hahn2011a} with the transfer function from \cite{Eisenstein1998a} at a redshift of $z_{ini}= 80$. The $\Lambda$CDM cosmological simulation is run using {GADGET-2} (\citealt{Springel2005} and \citealt{Springel2001}) \hl{is similar to the ones used in }\cite{Ramachandra2017}. The cosmological parameters used in the simulation are $\Omega_{m}= 0.276$, $\Omega_{\Lambda}= 0.724$, the Hubble parameter, $h = 0.703$, the power spectrum normalization, $\sigma_8 = 0.811$ and the spectral index $n_s= 0.961$.

% \begin{table}
%   \caption{Parameters for the simulation boxes: Side length $L$, number of particles $N_p$, mass of each particle $m_p$, and the gravitational softening length $\epsilon$ for the GADGET simulations are shown.}
% \begin{tabular}{l|c|c|c| }
% \hline
% $L$ & $N_p$ & $m_p$ & $\epsilon$  \\  \hline
% $100 h^{-1} Mpc$ & $128^3$      & $ 3.65 \times 10^{10} h^{-1} M_{\sun}$  & $20 h^{-1} kpc$\\ \hline
% $100 h^{-1} Mpc$ & $256^3$      & $ 4.57 \times 10^{9} h^{-1} M_{\sun}$  & $10 h^{-1} kpc$ \\ \hline
% 
% \end{tabular}
% \label{tab:Simulations}
% \end{table}

\hl{Multistream field $n_{str}(\mathbf{x})$ is calculated on the GADGET-2 snapshots at $z=0$ using the tessellation scheme by} \cite{Shandarin2012}. \hl{The multistream field can be computed at the native resolution of the Lagrangian grid of the simulation, i.e., at refinement factor of $l_l/l_d = 1$ (where $l_l$ is the inter-particle separation in Lagrangian grid and $l_d$ is the side length resolution of diagnostic grid). Arbitrarily high refinement factors can be utilized in computing multistream fields as well, for example  $l_l/l_d = 8$ for the halo multistream environment shown in \autoref{fig:full}. For analysis of full simulation boxes, we restrict $l_l/l_d$ to $1$ and $2$. }

\hl{Two halo finders are also used to identify potential haloes with 20 or more particles at $z= 0$: a classic Friends-of-Friends method} (FOF-\citealt{Davis1985}) using a popular linking length, $ b= 0.2$ (e.g. \citealt{Frenk1988} and \citealt{Lacey1994}) \hl{and the Adaptive Mesh Investigations of Galaxy Assembly }(AMIGA halo finder or AHF-\citealt{Knollmann2009a}, \citealt{Gill2004a}). \hl{Halo catalogue from these halo finders are used to compare with our implementation of halo detection in the multistream field. The haloes candidates from AHF and FOF algorithms are hereafter referred to as AHF-haloes and FOF-haloes respectively.}

@@

We intend to identify haloes in the $n_{str}(\bmath{x})$ field instead of using just the position coordinate data. While the eigenvalue analysis itself is done at a chosen time, the multistream field inherently has data from six-dimensional Lagrangian space $(\bmath{q}, \bmath{x})$ that contains the full dynamical information,similar to the phase-space sheet albeit in a different form. Dynamical history that is embedded in the multistream field is crucial in understanding the strength of gravitational binding of the particles. 
A physically motivated distinction between void and gravitationally collapsed regions \hl{-- voids are the regions with a single stream --} is a unique feature of multistream analysis (\citealt{Shandarin2012} and \citealt{Ramachandra2017}). Thus the haloes detected from local maxima of the $n_{str}$ field can be ensured to be away from the mono-streaming voids. Methods based on linking-length or density fields may not be able to ensure that  all the particles in haloes are away from voids (as shown for FOF haloes in \citealt{Ramachandra2017}) .

@@

\hl{New sub-section}

Hessian matrix $\mathbfss{H}(f)$ of a scalar field $f$ involves local second-order variations in three orthogonal directions. Each element of the Hessian matrix $\mathbfss{H}_{ij}(f)$ (where $i$ and $j$ can be any of $x$, $y$ or $z$ directions) is given by \autoref{eq:Hess}.

\begin{equation}
\label{eq:Hess}
 \mathbfss{H}_{ij}(f) = \frac{\partial^2 f}{\partial x_i \partial x_j}
\end{equation}

@@

(i) Every element of Hessian matrices $\mathbfss{H}(-n_{str})$, and consequently the eigenvalues $\lambda_i$'s are zero in single-stream voids. Even if the multistream field is a smoothed, the eigenvalues peak at zero. This property is unique to multistream fields. Eigenvalues of Hessians of density, velocity shear tensor \cite{Libeskind2013} do not peak at zero, and the eigenvalues of deformation tensor are negative in voids (by Zel’dovich formalism).

@@

{Our goal is to isolate the locations of convex geometries in the multistream flow field. Prospective regions of the halo locations in the web structure are selected by positive definite condition of the Hessian $\mathbfss{H}(-n_{str})$: $\lambda_1 > 0$, $\lambda_2 > 0$ and  $\lambda_3 > 0$, or simply the smallest of the eigenvalues, $\lambda_3 > 0$.
%\ghl{It should be stated explicitely that the above condition comes from the requirement of the boundary of the halo is convex at least in the 4-dimensional space (x,y,z,-nstr). Discuss the relation of the convexity in 4-space with that in 3-space (x,y,z) which is our principal def. of a halo - DONE in previous section . }
 We also filter out the regions if the multistream values inside do not suggest gravitational collapse into haloes. The sequence of our halo detection framework is listed below:

@@

\item Second order variations of the smoothed $-n_{str}(\bmath{x})$ is computed at each point in the field. This gives symmetric Hessian matrices for this field whose eigenvalues are real. Ordered eigenvalues of the Hessian, $ \lambda_1 \geq  \lambda_2 \geq \lambda_3$ are calculated. The $\lambda_3$ field is shown in the top right panel of \autoref{fig:labelsfilter}.

@@

\begin{figure}
\begin{minipage}[t]{0.99\linewidth}
 \centering\includegraphics[height=8.5cm]{fig3.pdf} 
\end{minipage}\hfill
\caption{ Detection of potential halo candidates in the multistream field: algorithm of segmentation and filtering are illustrated in a smaller slice of $40 h^{-1} Mpc \times 40 h^{-1} Mpc$ slice of the simulation box. Top left figure shows the multistream field of the slice. Voids (white) are the regions with $n_{str} =1$, rest are non-void structures. Blue patches within the structure (gray) are the regions with gravitational collapses in more than one direction, i.e., $n_{str} \geq 7$. Top Right figure shows the smallest eigenvalue $\lambda_3$ field. The value of $\lambda_3$ is close to 0 in most of the regions (yellow), including the voids. Regions with  $\lambda_3 > 0$ and $n_{str} > 1$, are isolated (black spots) using image segmentation techniques. Bottom left panel shows the filtering scheme: the red patches do not have maxima of $n_{str} \geq 7$ in the regions, hence are filtered out. The remaining potential halo regions with more than 20 particles are shown in the bottom right panel.}
\label{fig:labelsfilter}
\end{figure}

For the illustration halo detection framework in this section, we have calculated the number-of-streams at refinement factor of $2$ and smoothing scale of $0.39 h^{-1} Mpc$ (equal to the grid length of the multistream field) for the simulation box of $128^3$ particles and size $L = 100 h^{-1} Mpc$ . \hl{Hessian matrices and eigenvalues} are calculated on the same diagnosis grid. Results of the halo detection scheme for simulation box of higher mass resolution, and different smoothing factors are discussed in Sections \ref{sub:compareHalo} and \ref{sub:Smooth}. Hereafter we refer to the potential dark matter haloes detected from the Hessian analysis of the multistream field as $\lambda_3$-haloes for brevity.

>>>>

The web-like distribution of matter initially revealed by redshift surveys (with less than 300 galaxies by \citealt{Gregory1978} and around 1000 galaxies by \citealt{DeLapparent1986}) and numerical modeling (using N-body simulations of around 30000 particles by \citealt{Shandarin1983} and \citealt{Klypin1983a}) pioneered morphological investigations of the cosmic web structures (see \citealt{Bond1996}, also reviews by \citealt{Shandarin1989} and \citealt{Weygaert2008}). Detailed mapping of the Universe has crossed three million objects today, by catalogues such as the Sloan Digital Sky Survey (SDSS; \citealt{Albareti2016}). The upcoming Large Synoptic Survey Telescope (LSST; \citealt{lsst2009}) is expected to probe the nature of dark matter using several billion galaxies. On the other hand, cosmological simulations have improved immensely in several aspects -- numerical techniques, parallelization schemes, inclusion of various physical processes, volume and resolution (some of these developments are summarized in \citealt{Bertschinger1998} and \citealt{Bagla1997}). Modern state-of-the-art simulations like the Illustris Project \cite{Vogelsberger2014} and Q-Continuum \cite{Heitmann2015} use more than a billion dark matter particles. Finally, the ever improving data analysis techniques have resulted in new and sophisticated density estimators, geometrical and topological indicators. A plethora of algorithms for identifying and characterizing dark matter structures have emerged in last two decades (a summary on cosmological data analysis is highlighted in \citealt{Weygaert2009a}). Considering all these improvements, it is worth noting that the proto-structures detected in the modern simulations are qualitatively similar to the quasi-linear description of clustering by Zeldovich Approximation (ZA; \citealt{Zeldovich1970}). Location and properties of these structures, i.e., the voids, walls, filaments and haloes maybe inconsistent across different structure finding algorithms, but that is primarily due to varied definitions.

@@

Most of structure finders are halo finders only and majority of them are stemmed from three underlying algorithms. One of them is the SO (Spherical Over-density) halo finder that defines halos  as spherical regions whose mass density exceeds the mean density by a specified factor \citep{Press1974}. 
Another is the FOF  (Friends-of-Friends) halo finder describing haloes  as the groups of particles separated less than a specified linking length often chosen as 0.2 times the mean particle separation \citep{Davis1985}. The FOF can be also used for identifying filaments and walls/pancakes by increasing the linking length (\citealt{Zeldovich1982}, \citealt{Shandarin1983a}, \citealt{Shandarin2010b} )  %(refs in tex file and in additional\_refs.bib)}. 
%\citep{Zeldovich_etal:82,Shandarin:83,Shandarin_etal:10 } 
Finally the DENMAX (DENsity MAXimum) halo finder assumes that the halos are the peaks of the density  fields and thus selects the particles concentrated in the vicinity of the density maxima \citep{Bertschinger1991}. One of the common features of these  techniques is that all three are based on density, in one form or another. And all of them depend on free parameters that are chosen chiefly on the `merits principle' \citep{Forero-Romero2009a} rather than on physics. Over the years all three kinds of the halo finders have been experiencing various modifications and improvements.  A few examples from a long list of these modifications may include:

@@

 A detailed comparisons of several halo/sub-halo finders is provided in 
%done in projects like 
\cite{Knebe2011a}, \cite{Knebe2013}, \cite{Onions2012}. In a nice summary discussing  these developments as well as describing a few new suggestions they  concluded that there was no general consensus for a precise definition of a halo or a sub-halo. Consequently, there were different estimates of number of haloes, halo mass functions, halo centre locations, boundaries and other parameters.
%A  are given in comparison project papers quoted above.

@@

For FOF, the free parameter of linking length is generally taken as $b = 0.2$ times the mean separation of particles at $z=0$. This inter-particle separation corresponds to $\Delta_{vir} \approx 180$ if the halo has an isothermal density profile, $\rho \propto r^{-2}$. Using percolation theory, \cite{More2011} argued that this corresponds to a rather wide range of over-densities depending on halo mass and density profiles. They found out that $b = 0.2$ corresponds to local over-density $\delta$ within the enclosed halo to be in the range of 250 to 600. Moreover, the resulting FOF-haloes need not have a compact geometry: often the haloes are irregularly shaped, which is unlikely if the halos are virialized.
Hence modern algorithms re-define the halo boundaries by excluding particles using post-processing techniques. In recent simulations with clear delineation of walls and filaments \cite{Angulo2013a}, $b=0.05$ was used for finding FOF-haloes since the traditional value of $b = 0.2$ corresponded to structures that percolate into the web structure.

Absence of dynamical traits in the FOF and SO algorithms are arguably more crucial. In phase-space, the halo collapse models show collisionless DM particles in oscillatory motions about a core, at successive foldings of the phase-space sheet. The velocity field within each oscillatory spiral is multi-valued in physical space. Incrementing {\it multistream} shells, separated by {\it caustic} surfaces sequentially trace the structures of the cosmic web - walls, filaments and the haloes. Majority of the mass accretion into the haloes along the filaments: from lower multistreams into higher. Thus the DM haloes are not independent of filaments around them, and the hierarchical layers of multistreams represents this portrait precisely. This picture of structure formation was initially theorized using ZA \cite{Zeldovich1970} and in context of caustics \cite{Arnold1982} as well as in the Adhesion Approximation (\citealt{Gurbatov1989}, \citealt{Kofman1992}). 
\cite{Shandarin1989} reviewed gravitational evolution of density perturbations in the context.

It has been demonstrated that the multistream field in Eulerian space can be computed directly from the Lagrangian sub-manifold (\citealt{Shandarin2012} and \citealt{Abel2012b}). About 90\% of the field is single-streaming voids, and the rest of the volume comprises of multistream walls, filaments and haloes. \cite{Ramachandra2015} found the multistream value of $n_{str} \approx 90$ corresponds to virial density $\Delta_{vir} = 200$. On the other hand, DM particles are identified by \citep{Falck2012} as belonging to haloes if they undergo flip-flop along 3 orthogonal axes. These analyses have opened up a new avenue in studies of halo formation, both qualitative and quantitative. Re-investigations of halo spins, physical radii of the halo, sub-structure in the light of streaming phenomena have shown that the halo structures and formations are more complicated than previously believed. \cite{Vogelsberger2011} investigated the distribution of streams in small haloes at various redshifts. They concluded that tracking caustics and streams is better than density, since density fields are noisy in the dense inner regions of haloes. In another study, \cite{More2015} argued that the `splashback radius' - distance from the halo core to the first caustic enumerated from outside - is a better physical indicator of DM halo boundary than the virial radius (also see the discussion on turn-around radius of bound objects by \citealt{Lee2016a}). \cite{Angulo2013a} also agree with the view that the locally overdense regions correspond better with the volumes within the first caustic than the virialized DM clumps. Recent toy model of anisotropic halo collapse by \cite{Neyrinck2016} considers intersecting multistream filaments forming spinning nodes. \cite{Ramachandra2017} showed that the virial surfaces of FOF haloes have varying number of streams, including single-streams. Study by \cite{Shandarin2016} delineated the rich sub-structure of haloes using another derivative parameter from the Lagrangian sub-manifold called the `flip-flop' defined on the Lagrangian space.

@@

%\hl{New section}

\begin{figure*} 
\centering\includegraphics[height=7cm]{Chapter5/Source_v2/fig1.pdf} 
\caption{Dynamical collapse of dark matter in one-dimensional universe: Top panels show the $(\mathbf{p}, \mathbf{x})$ phase-space manifold of the dark matter sheet at redshifts $z_1$, $z_2$, $z_3$ and $z = 0$. Dots represent the dark matter particles. The momentum values are chosen at arbitrary scales. Bottom panels show the corresponding multistream field multistream field $n_{str}(\mathbf{x}, z)$ (red) and density field $\rho(\mathbf{x}, z)$ (gray). }
\label{fig:1d}
\end{figure*}

We begin with a simple illustration showing the formation of a few haloes in a one-dimensional simulation. Dark matter clustering in a (1+1)-dimensional phase-space $(\mathbf{p}, \mathbf{x})$ (where $p$ is the momentum and $x$ is the co-moving Eulerian coordinate) at four successive time steps is shown in the top panels of \autoref{fig:1d}. The lower panels show the corresponding multistream field (\citealt{Shandarin2012} and \citealt{Abel2012b}) $n_{str}(\mathbf{x}, z)$ (red) and density field $\rho(\mathbf{x}, z)$ (gray). At $z_1$ (left-most panel), velocity is single-valued in Eulerian co-ordinates shown, except at a small three-stream region near $\mathbf{x} = 5\pi/4$. This is the first instance of multistreaming in the region, which was previously had $n_{str} = 1$ throughout. The interface of $n_{str} = 1$ and $n_{str} = 3$ regions is also the location of the first caustic. On the other hand, the density calculated at a high resolution shows variations, even in the mono-streaming regions. The variations are especially more pronounced around  the caustic (near $\mathbf{x} = 5\pi/4$).

@@

\begin{figure} 
\centering\includegraphics[width=8cm]{Chapter5/Source_v2/fig2.pdf} 
\caption{Multistream field contours: The multistream field is calculated at 8 times the native resolution. Each inner convex blobs (red) surround local multistream maxima inside. Surrounding outer shell(blue) is not convex throughout the surface, and the outermost gray multistream surface displays a filamental geometry.}
\label{fig:full}
\end{figure}

@@

The emphasis of this paper is to demonstrate the use of multistream field in identifying potential dark matter haloes, and not a full statistical analysis of halo properties. For this purpose, we have run simulations at two different mass resolutions (number of particles $N_p = 128^3$ and $256^3$, and respective mass of particles, $m_p = 3.65 \times 10^{10} h^{-1} M_{\odot}$ and $ 4.57 \times 10^{9} h^{-1} M_{\odot}$), with the same periodic side length $L = 100 h^{-1} Mpc$. The gravitational softening length $\epsilon = 20 h^{-1} kpc$ and $10 h^{-1} kpc$ for low and high resolution simulation respectively. The initial conditions are generated by {MUSIC} \citep{Hahn2011a} with the transfer function from \cite{Eisenstein1998a} at a redshift of $z_{ini}= 80$. The $\Lambda$CDM cosmological simulation is run using {GADGET-2} (\citealt{Springel2005} and \citealt{Springel2001}) is similar to the ones used in \cite{Ramachandra2017}. The cosmological parameters used in the simulation are $\Omega_{m}= 0.276$, $\Omega_{\Lambda}= 0.724$, the Hubble parameter, $h = 0.703$, the power spectrum normalization, $\sigma_8 = 0.811$ and the spectral index $n_s= 0.961$.

% \begin{table}
%   \caption{Parameters for the simulation boxes: Side length $L$, number of particles $N_p$, mass of each particle $m_p$, and the gravitational softening length $\epsilon$ for the GADGET simulations are shown.}
% \begin{tabular}{l|c|c|c| }
% \hline
% $L$ & $N_p$ & $m_p$ & $\epsilon$  \\  \hline
% $100 h^{-1} Mpc$ & $128^3$      & $ 3.65 \times 10^{10} h^{-1} M_{\odot}$  & $20 h^{-1} kpc$\\ \hline
% $100 h^{-1} Mpc$ & $256^3$      & $ 4.57 \times 10^{9} h^{-1} M_{\odot}$  & $10 h^{-1} kpc$ \\ \hline
% 
% \end{tabular}
% \label{tab:Simulations}
% \end{table}

Multistream field $n_{str}(\mathbf{x})$ is calculated on the GADGET-2 snapshots at $z=0$ using the tessellation scheme by \cite{Shandarin2012}. The multistream field can be computed at the native resolution of the Lagrangian grid of the simulation, i.e., at refinement factor of $l_l/l_d = 1$ (where $l_l$ is the inter-particle separation in Lagrangian grid and $l_d$ is the side length resolution of diagnostic grid). Arbitrarily high refinement factors can be utilized in computing multistream fields as well, for example  $l_l/l_d = 8$ for the halo multistream environment shown in \autoref{fig:full}. For analysis of full simulation boxes, we restrict $l_l/l_d$ to $1$ and $2$.

Two halo finders are also used to identify potential haloes with 20 or more particles at $z= 0$: a classic Friends-of-Friends method (FOF-\citealt{Davis1985}) using a popular linking length, $ b= 0.2$ (e.g. \citealt{Frenk1988} and \citealt{Lacey1994}) and the Adaptive Mesh Investigations of Galaxy Assembly (AMIGA halo finder or AHF-\citealt{Knollmann2009a}, \citealt{Gill2004a}). Halo catalogue from these halo finders are used to compare with our implementation of halo detection in the multistream field. The haloes candidates from AHF and FOF algorithms are hereafter referred to as AHF-haloes and FOF-haloes respectively.

@@

We intend to identify haloes in the $n_{str}(\mathbf{x})$ field instead of using just the position coordinate data. While the eigenvalue analysis itself is done at a chosen time, the multistream field inherently has data from six-dimensional Lagrangian space $(\mathbf{q}, \mathbf{x})$ that contains the full dynamical information,similar to the phase-space sheet albeit in a different form. Dynamical history that is embedded in the multistream field is crucial in understanding the strength of gravitational binding of the particles. 
A physically motivated distinction between void and gravitationally collapsed regions -- voids are the regions with a single stream -- is a unique feature of multistream analysis (\citealt{Shandarin2012} and \citealt{Ramachandra2017}). Thus the haloes detected from local maxima of the $n_{str}$ field can be ensured to be away from the mono-streaming voids. Methods based on linking-length or density fields may not be able to ensure that  all the particles in haloes are away from voids (as shown for FOF haloes in \citealt{Ramachandra2017}) .

@@

%\hl{New sub-section}

Hessian matrix $\mathbf{H}(f)$ of a scalar field $f$ involves local second-order variations in three orthogonal directions. Each element of the Hessian matrix $\mathbf{H}_{ij}(f)$ (where $i$ and $j$ can be any of $x$, $y$ or $z$ directions) is given by \autoref{eq:Hess}.

\begin{equation}
\label{eq:Hess}
 \mathbf{H}_{ij}(f) = \frac{\partial^2 f}{\partial x_i \partial x_j}
\end{equation}

@@

(i) Every element of Hessian matrices $\mathbf{H}(-n_{str})$, and consequently the eigenvalues $\lambda_i$'s are zero in single-stream voids. Even if the multistream field is a smoothed, the eigenvalues peak at zero. This property is unique to multistream fields. Eigenvalues of Hessians of density, velocity shear tensor \cite{Libeskind2013} do not peak at zero, and the eigenvalues of deformation tensor are negative in voids (by Zel’dovich formalism).

@@

{Our goal is to isolate the locations of convex geometries in the multistream flow field. Prospective regions of the halo locations in the web structure are selected by positive definite condition of the Hessian $\mathbf{H}(-n_{str})$: $\lambda_1 > 0$, $\lambda_2 > 0$ and  $\lambda_3 > 0$, or simply the smallest of the eigenvalues, $\lambda_3 > 0$.
%\ghl{It should be stated explicitely that the above condition comes from the requirement of the boundary of the halo is convex at least in the 4-dimensional space (x,y,z,-nstr). Discuss the relation of the convexity in 4-space with that in 3-space (x,y,z) which is our principal def. of a halo - DONE in previous section . }
 We also filter out the regions if the multistream values inside do not suggest gravitational collapse into haloes. The sequence of our halo detection framework is listed below:

@@

\item Second order variations of the smoothed $-n_{str}(\mathbf{x})$ is computed at each point in the field. This gives symmetric Hessian matrices for this field whose eigenvalues are real. Ordered eigenvalues of the Hessian, $ \lambda_1 \geq  \lambda_2 \geq \lambda_3$ are calculated. The $\lambda_3$ field is shown in the top right panel of \autoref{fig:labelsfilter}.

@@

\begin{figure}
\begin{minipage}[t]{0.99\linewidth}
 \centering\includegraphics[height=8.5cm]{Chapter5/Source_v2/fig3.pdf} 
\end{minipage}\hfill
\caption{ Detection of potential halo candidates in the multistream field: algorithm of segmentation and filtering are illustrated in a smaller slice of $40 h^{-1} Mpc \times 40 h^{-1} Mpc$ slice of the simulation box. Top left figure shows the multistream field of the slice. Voids (white) are the regions with $n_{str} =1$, rest are non-void structures. Blue patches within the structure (gray) are the regions with gravitational collapses in more than one direction, i.e., $n_{str} \geq 7$. Top Right figure shows the smallest eigenvalue $\lambda_3$ field. The value of $\lambda_3$ is close to 0 in most of the regions (yellow), including the voids. Regions with  $\lambda_3 > 0$ and $n_{str} > 1$, are isolated (black spots) using image segmentation techniques. Bottom left panel shows the filtering scheme: the red patches do not have maxima of $n_{str} \geq 7$ in the regions, hence are filtered out. The remaining potential halo regions with more than 20 particles are shown in the bottom right panel.}
\label{fig:labelsfilter}
\end{figure}

For the illustration halo detection framework in this section, we have calculated the number-of-streams at refinement factor of $2$ and smoothing scale of $0.39 h^{-1} Mpc$ (equal to the grid length of the multistream field) for the simulation box of $128^3$ particles and size $L = 100 h^{-1} Mpc$ . Hessian matrices and eigenvalues are calculated on the same diagnosis grid. Results of the halo detection scheme for simulation box of higher mass resolution, and different smoothing factors are discussed in Sections \ref{sub:compareHalo} and \ref{sub:Smooth}. Hereafter we refer to the potential dark matter haloes detected from the Hessian analysis of the multistream field as $\lambda_3$-haloes for brevity.
